---
title: |
  | oyoyo
  | hogehoge研修
author: "iotsuka"
date: '2023-06-12'
output:
  beamer_presentation:
    theme: default
    latex_engine: lualatex
    keep_tex: true
    includes:
      in_header: header.tex
    fig_caption: true
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(
  dev = "cairo_pdf",
  dev.args = list(family = "Meiryo")
)
```


```{r include=FALSE}
library(tidyverse)
library(readxl)
library(viridis)
library(gt)
library(ragg)
library(ggthemes)
```


# a beautiful story

\Huge

\begin{itemize}
\item[\textbullet] 予算の会議\pause
\item[\textbullet] ある事業について\pause
\item[\textbullet] どんな成果?
\end{itemize}

# a beautiful story

\IfFileExists{kids.jpg}{\centering\includegraphics{kids.jpg}}{\relax}

# evidence-based policy-making
\Huge

\begin{itemize}
\item[\textbullet] 思いだけでは進まない
\pause
\item[\textbullet] 客観的な根拠
\end{itemize}
\pause
\vspace*{-30pt}
\hspace*{18pt}\rotatebox{13}{\scalebox{2}{\textcolor{gray}{dataの活用!}}}

# case study 1


# case study 1

\Huge


\rotatebox{0}{\scalebox{2}{\textcolor{gray}{\begin{tabular}{l}病気の検査\end{tabular}}}}

\vspace*{-48pt}
\pause
\rotatebox{20}{\scalebox{3.8}{\textcolor{red!30}{\begin{tabular}{l}\fbox{\fbox{陽性}}\end{tabular}}}}

<!-- # case study 1 -->

<!-- \Huge -->

<!-- 検査を受けました\pause{\textrightarrow{}\,\,{}陽性!!!} -->

# case study 1

\LARGE

\begin{itemize}
\item どんな病気??
\pause
\item 　\textbullet{}\hspace{2pt}人口の1\%がこの病気
\pause
\item
\item どんな検査??
\pause
\item 　\textbullet{}\hspace{2pt}病気なら90\%の確率で陽性
\pause
\item 　\textbullet{}\hspace{2pt}病気でないなら90\%の確率で陰性
\end{itemize}

# case study 1

\Huge

実際に病気である確率は??



```{r}
prevalence <- 1    #有病率
sensitivity <-  90  #感度
specificity <-  90  #特異度

precision <- function(x, y, z){
  x * y / (x * y + (100 - x) * (100 - z))
}

accuracy <- function(x, y, z){
  (x * y + (100 - x) * z) / 10000
}
```






# case study 1

\scalebox{.88}{%
\begin{tikzpicture}
 \tikzset{block/.style={rectangle, text width=25mm, text centered, rounded corners, minimum height=1.5cm}};
%\draw [help lines] (-6,0) grid (6,6);%(0,0)から(10,4)までの"細線の方眼"
\node[block,fill=gray!10] (population) at (0,7) {population\\1,000人};
\pause
\node[block,fill=gray!10] (sick) at (-3.5,4) {sick\\10人};
\draw [->] (population) to[out=-90, in=60] node[auto=right] {$1\%$} (sick); 
\pause
\node[block,fill=gray!10] (not_sick) at (3.5,4) {not sick\\990人};
\draw [->] (population) to[out=-90, in=120]  node[auto=left] {$99\%$} (not_sick); 
\pause
\node[block,fill=red!10] (true_positive) at (-5,) {true positive\\9人};
\draw [->] (sick) to[out=-90, in=90]  node[auto=right] {$90\%$} (true_positive); 
\pause
\node[block,fill=gray!10] (false_negative) at (-2,1) {false negative\\1人};
\draw [->] (sick) to[out=-90, in=90]  node[auto=left] {$10\%$} (false_negative); 
\pause
\node[block,fill=gray!10] (true_negative) at (2,1) {true negative\\891人};
\draw [->] (not_sick) to[out=-90, in=90]  node[auto=right] {$90\%$} (true_negative); 
\pause
\node[block,fill=red!10] (false_positive) at (5,1) {false positive\\99人};
\draw [->] (not_sick) to[out=-90, in=90]  node[auto=left] {$10\%$} (false_positive); 
 \end{tikzpicture}%
}

# case study 1
\begin{center}\Huge
\[
\frac{9}{9 + 99}=0.083333
\ldots
\]
\end{center}




# case study 2 --- world war II

\raggedleft
\IfFileExists{survivorship_bias.png}{\scalebox{.9}{\includegraphics{survivorship_bias.png}}}{\relax}


\tiny

\raggedleft
Martin Grandjean(vector), McGeddon(picture), Cameron Moll(concept)

\vspace{-5pt}

This work is licensed under the Creative Commons Attribution-Share Alike 4.0 International license.



# case study 3

\Huge

\begin{itemize}
\item[\textbullet] ある町
\pause
\item[\textbullet] 4つの学校
\pause
\item[\textbullet] 同じテスト
\end{itemize}

# case study 3

\Huge

\begin{itemize}
\item[\textbullet] 家庭での学習時間
\pause
\item[\textbullet] 正答率
\end{itemize}
\pause
\vspace*{-55pt}
\rotatebox{20}{\scalebox{2}{\textcolor{gray}{どんな関係?}}}


# plot

```{r echo=FALSE}
df_base <- iris %>%
  janitor::clean_names() %>%
  as_tibble() %>% 
  filter(species == "versicolor") %>% 
  rename("time" = sepal_length,
         "point" = petal_length,
         "school" = species)
```

```{r echo=FALSE}
df_D <- df_base %>% 
  mutate(time = time - 3, point = point + .4, school = "school_D")

df_C <- df_D %>% 
  mutate(time = time -.25, point = point + .9, school = "school_C")

df_B <- df_C %>% 
  mutate(time = time -.25, point = point + .9, school = "school_B")

df_A <- df_B %>% 
  mutate(time = time -.25, point = point + .9, school = "school_A")
```




```{r echo=FALSE}
q <- df_A %>% bind_rows(df_B, df_C, df_D) %>% 
  ggplot(aes(time, point * 10)) +
  geom_jitter(width = .8, height = .8, alpha = .7, size = 2) +
#  geom_point(alpha = .75, size = 2) +
  labs(x = "time spent studying", 
       y = "score")
```


```{r}
set.seed(1)
q
```


# case study 3



```{r echo=FALSE, message=FALSE}
set.seed(1)
q +
  geom_smooth(method = "lm", se = FALSE)
```


# case study 3

\IfFileExists{./letsnotsee.jpg}{\includegraphics[width=\textwidth]{./letsnotsee.jpg}}{\relax}


# case study 3



```{r echo=FALSE}
set.seed(1)
q +
  aes(colour = school) +
#  scale_color_viridis_d(name = "") 
  ggthemes::scale_colour_few(guide = "none")
```




# case study 3



```{r echo=FALSE}
set.seed(1)
q +
  aes(colour = school) +
  scale_color_few(guide = "none") +
  facet_wrap(~school)
```

# case study 3



```{r echo=FALSE}
set.seed(1)
q +
  aes(colour = school) +
  scale_color_few(guide = "none") +
  facet_wrap(~school) +
  geom_smooth(method = "lm", se = FALSE)
```

# case study 3



```{r echo=FALSE}
set.seed(1)
q +
  aes(colour = school) +
  scale_color_few(guide = "none") +
  geom_smooth(method = "lm", se = FALSE)
```

# case study 3



```{r echo=FALSE, message=FALSE}
set.seed(1)
q +
  geom_smooth(method = "lm", se = FALSE)
```


# case study 3
\LARGE

\begin{itemize}
\item[\textbullet] 全体では\par
「勉強するほど成績がさがる」
\bigskip\pause
\item[\textbullet] 学校ごとでは\par
「勉強するほど成績があがる」
\end{itemize}

# case study 3 --- Simpson's Paradox
\LARGE

全体と個々のグループでの傾向が異なる

\normalsize

# tebiki

\vspace*{-20pt}
\IfFileExists{./tebiki.jpg}{\centering\scalebox{.9}{\rotatebox{-90}{\includegraphics[width=\textheight]{tebiki.jpg}}}}{\relax}


# Sir Winston Spencer Churchill

\raggedleft\Huge

\raisebox{100pt}{\rotatebox{30}{\textcolor{white}{\scalebox{1.2}{Brevity}}}}　　
\IfFileExists{Sir_Winston_Churchil.jpg}{\scalebox{.75}{\includegraphics{Sir_Winston_Churchil.jpg}}}{\relax}


\tiny

\raggedleft
The Roaring Lion(Yousuf Karsh)
\vspace{-5pt}

This work is licensed under the Creative Commons Attribution 2.0 Generic License.

# Sir Winston Spencer Churchill

\raggedleft\Huge

\raisebox{100pt}{\rotatebox{30}{\textcolor{black}{\scalebox{1.2}{Brevity}}}}　　
\IfFileExists{Sir_Winston_Churchil.jpg}{\scalebox{.75}{\includegraphics{Sir_Winston_Churchil.jpg}}}{\relax}


\tiny

\raggedleft
The Roaring Lion(Yousuf Karsh)
\vspace{-5pt}

This work is licensed under the Creative Commons Attribution 2.0 Generic License.




# Brevity

\centering
\vspace*{-21.5pt}
\IfFileExists{churchill_memo.jpg}{\scalebox{1.025}{\includegraphics{churchill_memo.jpg}}}{\relax}


# Brevity
\Large

To do our work, we all have to read a mass of papers. 
Nearly all of them are far too long.\par

\vfill
\Huge

\rotatebox{15}{\scalebox{1.6}{\textcolor{white}{\begin{tabular}{l}大量の書類\\長すぎる\end{tabular}}}}

\vfill

# Brevity
\Large

To do our work, we all have to read \textcolor{olive}{a mass of papers}. 
Nearly all of them are \textcolor{olive}{far too long}.\par


\vfill
\Huge

\rotatebox{15}{\scalebox{1.6}{\textcolor{olive}{\begin{tabular}{l}大量の書類\\長すぎる\end{tabular}}}}

\vfill

# Brevity
\Large


This wastes time, 
while energy has to be spent in looking for the essential points.\par

\vfill\Huge

\rotatebox{15}{\scalebox{1.6}{\textcolor{white}{\begin{tabular}{l}時間のむだ\\要点がわからない\end{tabular}}}}

\vfill
\mbox{}

# Brevity
\Large


This \textcolor{olive}{wastes time}, 
while energy has to be spent in looking for \textcolor{olive}{the essential points}.\par

\vfill\Huge

\rotatebox{15}{\scalebox{1.6}{\textcolor{olive}{\begin{tabular}{l}時間のむだ\\要点がわからない\end{tabular}}}}

\vfill


# Brevity
\Large


I ask my colleagues and their staffs to see to it that their Reports are shorter.


\vfill\Huge

\rotatebox{15}{\scalebox{1.6}{\textcolor{white}{\begin{tabular}{l}報告書を短く\end{tabular}}}}

\vfill

# Brevity
\Large


I ask my colleagues and their staffs to see to it that \textcolor{olive}{their Reports are shorter}.


\vfill\Huge

\rotatebox{15}{\scalebox{1.6}{\textcolor{olive}{\begin{tabular}{l}報告書を短く\end{tabular}}}}

\vfill


# Brevity
\Huge



\begin{itemize}
\item \scalebox{1.3}{歯切れよくいいきれ}
  \begin{itemize}
  \item \pause
  \item \scalebox{2}{\textbullet{}\myCrossedOut{red!60!gray}{冗長ないいまわし}}
  \item \scalebox{2}{\textbullet{}乱暴な表現でよし}
\end{itemize}
\end{itemize}


# what is important

\Huge

-  \textbullet{}\hspace{2pt} 本質的なこと
\bigskip
-  \textbullet{}\hspace{2pt} 体裁
\bigskip
-  \textbullet{}\hspace{2pt} スピード感



# tebiki

\vspace*{-20pt}
\IfFileExists{./tebiki.jpg}{\centering\scalebox{.9}{\rotatebox{-90}{\includegraphics[width=\textheight]{tebiki.jpg}}}}{\relax}


# quiz

\Huge

- 「又は」or「または」
\bigskip
- 「子供」or「子ども」
\bigskip
- 「手引」or「手引き」

# manual

\IfFileExists{./seisaku_homu01.png}{\scalebox{1}{\includegraphics{seisaku_homu01.png}}}{\relax}

# manual

\IfFileExists{./seisaku_homu03.png}{\scalebox{1.335}{\includegraphics{seisaku_homu03.png}}}{\relax}



# tebiki

\vspace*{-20pt}
\IfFileExists{./tebiki.jpg}{\centering\scalebox{.9}{\rotatebox{-90}{\includegraphics[width=\textheight]{tebiki.jpg}}}}{\relax}



# attraction



\vspace*{-4pt}
\IfFileExists{./bento.jpg}{\centering\scalebox{.85}{\includegraphics[width=\textheight]{bento.jpg}}}{\relax}



# last but not least



# data

- 4つのデータ(data\_1 ～ data\_4)があります

- それぞれのデータは、11組のxとyから構成されています

```{r echo=FALSE}
df_1 <- anscombe %>%
  as_tibble() %>% 
  select(x1, y1) %>% 
  rename("x" =x1, "y" = y1)

df_2 <- anscombe %>%
  as_tibble() %>% 
  select(x2, y2) %>% 
  rename("x" =x2, "y" = y2)

df_3 <- anscombe %>%
  as_tibble() %>% 
  select(x3, y3) %>% 
  rename("x" =x3, "y" = y3)

df_4 <- anscombe %>%
  as_tibble() %>% 
  select(x4, y4) %>% 
  rename("x" =x4, "y" = y4)
```


# data

\begin{tabular}{rr}
\multicolumn{2}{l}{data\_1}\\
\toprule
x & y\\
\midrule
10 & 8.04\\
8 & 6.95\\
13 & 7.58\\
9 & 8.81\\
11 & 8.33\\
14 & 9.96\\
6 & 7.24\\
4 & 4.26\\
12 & 10.84\\
7 & 4.82\\
5 & 5.68\\
\bottomrule
\end{tabular}\hfill
\begin{tabular}{rr}
\multicolumn{2}{l}{data\_2}\\
\toprule
x & y\\
\midrule
10 & 9.14\\
8 & 8.14\\
13 & 8.74\\
9 & 8.77\\
11 & 9.26\\
14 & 8.10\\
6 & 6.13\\
4 & 3.10\\
12 & 9.13\\
7 & 7.26\\
5 & 4.74\\
\bottomrule
\end{tabular}\hfill
\begin{tabular}{rr}
\multicolumn{2}{l}{data\_3}\\
\toprule
x & y\\
\midrule
10 & 7.46\\
8 & 6.77\\
13 & 12.74\\
9 & 7.11\\
11 & 7.81\\
14 & 8.84\\
6 & 6.08\\
4 & 5.39\\
12 & 8.15\\
7 & 6.42\\
5 & 5.73\\
\bottomrule
\end{tabular}\hfill
\begin{tabular}{rr}
\multicolumn{2}{l}{data\_4}\\
\toprule
x & y\\
\midrule
8 & 6.58\\
8 & 5.76\\
8 & 7.71\\
8 & 8.84\\
8 & 8.47\\
8 & 7.04\\
8 & 5.25\\
19 & 12.50\\
8 & 5.56\\
8 & 7.91\\
8 & 6.89\\
\bottomrule
\end{tabular}



# data

それぞれどういうデータなんだろう

- 表だけ見てもよくわからない

# data

それぞれどういうデータなんだろう

- xの平均
- xの分散
- yの平均
- yの分散
- xとyの相関係数
- 回帰直線の式

# summarise

```{r}
sokan_keisu <- list(df_1, df_2, df_3, df_4) %>% map_dbl(~cor(.x$x, .x$y))
list(df_1, df_2, df_3, df_4) %>% map_dfr(~summarise(.x, across(.cols=everything(), .fns=list(mean=mean, var=var), .names="{col}_{fn}"))) %>% 
  add_column(data=c("data_1", "data_2", "data_3", "data_4"), .before ="x_mean") %>% 
  add_column(cor = sokan_keisu, .after= "y_var") %>% 
  knitr::kable()
```

>- 要約統計量はよく似ている

# visualization

```{r fig.cap="`ggplot2` によるグラフ", echo=FALSE}
anscombe %>% 
  pivot_longer(cols = everything(),
               names_to = c(".value", "set"),
               names_pattern = "(x|y)(\\d)") %>% 
  mutate(set=str_c("data_", set)) %>% 
  ggplot(aes(x, y)) +
  geom_point() +
  facet_wrap(~set)
```



# visualization

```{r fig.cap="`ggplot2` によるグラフ", echo=FALSE}
anscombe %>% 
  pivot_longer(cols = everything(),
               names_to = c(".value", "set"),
               names_pattern = "(x|y)(\\d)") %>% 
  mutate(set=str_c("data_", set)) %>% 
  ggplot(aes(x, y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~set)
```


# lesson

表でみてもよくわからない

平均値とか分散を計算してもよくわからない

図示がだいじ


# quiz:How many millions are in a trillion?


```{r echo=FALSE}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width,
                      colour = Species)) +
  geom_point()
spc <- c("ヒオウギアヤメ", "イリス・ウェルシコロル",
         "イリス・ヴァージニカ")
p + labs(x = "がくの長さ(cm)", y = "萼の幅(cm)") +
         scale_colour_discrete(name = "種", labels = spc)
```




# quiz:How many millions are in a trillion?


\LARGE

1兆は100万の何倍でしょう？


\begin{tabular}{rr}\toprule
回答&率\\\midrule
千倍&18\%\\
万倍&12\%\\
10万倍&21\%\\
100万倍&21\%\\
1000万倍&17\%\\
わからない&12\%\\\bottomrule
\end{tabular}





# pie chart

```{r include=FALSE}
choices = c("10^3 倍","10^4 倍","10^5 倍","10^6 倍","10^7 倍","わからない") %>% fct_inorder() 
value <-  c(18,12,21,21,17,12)
df_pie <- tibble(choices, value)
```


```{r out.height="8cm"}
df_pie %>% 
　　mutate(position = 100 - (cumsum(value) - value/2)) %>% 
  ggplot(aes(x = "", y = value, fill = choices)) + 
　　geom_bar(stat = "identity", position = "stack") +
　　coord_polar(theta = "y", direction = -1) + 
　　geom_text(aes(y = position, label = choices), size = 6, colour = "white")  +
    scale_fill_viridis_d(guide = "none") +
    theme_void()
```


# substitute for a pie chart


```{r echo=FALSE}
df_pie %>% 
  ggplot() + geom_col(aes(choices, value, fill = choices))  +
  scale_fill_viridis_d(guide = "none") +
  labs(x = "", y = "比率(%)")
```


# substitute for a pie chart


```{r}
df_pie %>% 
  ggplot() + geom_col(aes(fct_rev(choices), value, fill = choices)) +
  scale_fill_viridis_d(guide = "none") +
  coord_flip() +
  labs(x = "", y = "比率(%)")
```


# substitute for a pie chart


```{r}
df_pie %>% 
  ggplot() + geom_point(aes(fct_rev(choices), value, colour = choices),
                        colour = "orange",
                        size = 10) +
  geom_segment(aes(x = choices, xend = choices,
                   y = 0, yend = value),
               colour = "grey",
               size = 1) +
  coord_flip() +
  labs(x = "", y = "比率(%)")
```



# quiz:How many millions are in a trillion?


```{r echo=FALSE}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width,
                      colour = Species)) +
  geom_point()
spc <- c("ヒオウギアヤメ", "イリス・ウェルシコロル",
         "イリス・ヴァージニカ")
p +  scale_colour_discrete(name = "種", labels = iris$Species) +
     labs(x = "ほげほげ(cm)", y = "およよ(cm)") 
```


# Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```





# 文書・資料


- \textbullet{}\hspace{2pt} The aim should be Reports which set out the main points in a series of short, crisp paragraphs.
-  \textbullet{}\hspace{2pt} If a Report relies on detailed analysis of some complicated factors, or on statistics, these should be set out in an Appendix.
-  \textbullet{}\hspace{2pt} Often the occasion is best met by submitting not a full-dress Report, but an Aide-memoire consisting of headings only, which can be expanded orally if needed. 
- \textbullet\hspace{2pt}Let us have an end of such phrases as these: 
 “It is also of importance to bear in mind the following considerations...”, or  “Consideration should be given to the possibility of carrying into effect...”
Most of these woolly phrases are mere padding, which can be left out altogether, or replaced by a single word. Let us not shrink from using the short expressive phrase, even if it is conversational.

# 文書・資料


Reports drawn up on the lines I propose may at first seem rough as compared with the flat surface of officialese jargon. But the saving in time will be great, while the discipline of setting out the real points concisely will prove an aid to clearer thinking.





# tohoho


```{r include=FALSE}
choices = c("ほぼ毎日","週2、3回","週1回","2週間に1回","月1回","学校による", "有事のみ", "未実施", "把握せず") %>% fct_inorder() 
value <-  c(11.3,5.7,17,7.5,13.2,3.8,3.8,34,1.9)
df_pie <- tibble(choices, value)
```


```{r}
school <- c("school_A", "school_B", "school_C")
とても不満 <- c(5, 18, 20)
不満 <- c(20, 18, 25)
ふつう <- c(13, 15, 30)
満足 <- c(43, 30, 15)
とても満足 <- c(17, 20, 10)

df_likert <- tibble(school, とても不満, 不満, ふつう, 満足, とても満足)
```

```{r}
df_likert %>%
  pivot_longer(-school) %>%
  mutate(name = fct_inorder(name)) %>% 
  mutate(name = fct_rev(name)) %>% 
  ggplot(data = .)+
  geom_bar(aes(x = fct_rev(school), y=value, fill=name), position="stack", stat="identity")+
  coord_flip() + 
  ggtitle("授業の満足度")+
  ylab("Percentage")+
  xlab("School")+
  scale_fill_brewer(palette="PRGn", name = "" ) +
  theme(legend.position="bottom") +
  guides(fill = guide_legend(reverse = TRUE))
```


# tohoho
 



```{r}
df_likert %>% HH::likert(school~., ., as.percent = TRUE,
           main="授業の満足度",
           xlab="percentage", ylab="School") 
```


## 誕生日問題

閏年は無視。
k人の集団で同じ誕生日の人がいるかどうか。

### 少なくとも1組はいる確率

```{r}
birthday <- function(k){
     {replicate(1000, {
         len <- sample(365, k, replace = TRUE) |> unique() |> length()
         k - len
     }) > 0} |> mean()
 }
```



```{r}
Vectorize(birthday)(1:40)
```

### 同じ誕生日の

```{r}
birthday_daburi <- function(k){
     {replicate(1000, {
         len <- sample(365, k, replace = TRUE) |> unique() |> length()
         k - len
     }) } |> mean()
 }
```


```{r}
Vectorize(birthday_daburi)(1:40)
```

