---
title: |
  | oyoyo
  | hogehoge研修
author: "iotsuka"
date: '2023-06-12'
output:
  beamer_presentation:
    theme: default
    latex_engine: lualatex
    keep_tex: true
    includes:
      in_header: header.tex
    fig_caption: true
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(
  dev = "cairo_pdf",
  dev.args = list(family = "Meiryo")
)
```


```{r include=FALSE}
library(tidyverse)
library(readxl)
library(viridis)
library(gt)
library(ragg)
```


# a beautiful story

\Huge

- 予算の会議
- ある事業について
- どんな成果?

# a beautiful story

\IfFileExists{kids.jpg}{\includegraphics{kids.jpg}}{\relax}

# evidence-based policy-making
\Huge

- 思いだけでは進まない
- dataを活用しよう

# evidence-based policy-making
\Huge

\begin{itemize}
\item \textbullet{}\hspace{2pt}思いだけでは進まない
\pause
\item \textbullet{}\hspace{2pt}dataを活用しよう
\end{itemize}


# quiz:whaddayathink?

\Huge



検査を受けました\pause{\textrightarrow{}\,\,{}陽性!!!}




# quiz:whaddayathink?

\LARGE

\begin{itemize}
\item どんな病気??
\pause
\item 　\textbullet{}\hspace{2pt}1\%の人がこの病気にかかってる
\pause
\item 
\item どんな検査??
\pause
\item 　\textbullet{}\hspace{2pt}病気なら90\%の確率で陽性
\pause
\item 　\textbullet{}\hspace{2pt}病気でないなら90\%の確率で陰性
\end{itemize}

# quiz:whaddayathink?

\Huge

実際に病気である確率は??



# answer

- 有病率1\%, 感度90\%, 特異度90\%

\bigskip
\bigskip

- \hspace*{-25pt}\scalebox{.7}{%
\begin{tikzpicture}
[
    level 1/.style={sibling distance=80mm},
    level 2/.style={sibling distance=60mm},
]
	\node {1,000人}
		child {
        node {\begin{tabular}{c}病気\\10人\end{tabular}}
		      child {node {%
                       \begin{tabular}{c}
                       9人\\(真陽性)
                       \end{tabular} %
                       }
              }
		      child {node {%
		                   \begin{tabular}{c}
                       1人\\(偽陰性)
                       \end{tabular}%
                       }
                 }
            }
		child {
		    node {\begin{tabular}{c}病気でない\\990人\end{tabular}}
		      child {node {%
		                   \begin{tabular}{c}
                       891人\\(真陰性)
                       \end{tabular}%
            	  	     }
              	}
		      child {node {%
                       \begin{tabular}{c}
                       99人\\(偽陽性)
                       \end{tabular}%
                       }
                }
     	  	  };
\end{tikzpicture}%
}

# answer
\begin{center}\Huge
$\frac{9}{9 + 99}=0.083333$
\end{center}

```{r}
prevalence <- .8    #有病率
sensitivity <-  90  #感度
specificity <-  93  #特異度

precision <- function(x, y, z){
  x * y / (x * y + (100 - x) * (100 - z))
}

accuracy <- function(x, y, z){
  (x * y + (100 - x) * z) / 10000
}
```













# data

- 4つのデータ(data\_1 ～ data\_4)があります

- それぞれのデータは、11組のxとyから構成されています

```{r echo=FALSE}
df_1 <- anscombe %>%
  as_tibble() %>% 
  select(x1, y1) %>% 
  rename("x" =x1, "y" = y1)

df_2 <- anscombe %>%
  as_tibble() %>% 
  select(x2, y2) %>% 
  rename("x" =x2, "y" = y2)

df_3 <- anscombe %>%
  as_tibble() %>% 
  select(x3, y3) %>% 
  rename("x" =x3, "y" = y3)

df_4 <- anscombe %>%
  as_tibble() %>% 
  select(x4, y4) %>% 
  rename("x" =x4, "y" = y4)
```


# data

\begin{tabular}{rr}
\multicolumn{2}{l}{data\_1}\\
\toprule
x & y\\
\midrule
10 & 8.04\\
8 & 6.95\\
13 & 7.58\\
9 & 8.81\\
11 & 8.33\\
14 & 9.96\\
6 & 7.24\\
4 & 4.26\\
12 & 10.84\\
7 & 4.82\\
5 & 5.68\\
\bottomrule
\end{tabular}\hfill
\begin{tabular}{rr}
\multicolumn{2}{l}{data\_2}\\
\toprule
x & y\\
\midrule
10 & 9.14\\
8 & 8.14\\
13 & 8.74\\
9 & 8.77\\
11 & 9.26\\
14 & 8.10\\
6 & 6.13\\
4 & 3.10\\
12 & 9.13\\
7 & 7.26\\
5 & 4.74\\
\bottomrule
\end{tabular}\hfill
\begin{tabular}{rr}
\multicolumn{2}{l}{data\_3}\\
\toprule
x & y\\
\midrule
10 & 7.46\\
8 & 6.77\\
13 & 12.74\\
9 & 7.11\\
11 & 7.81\\
14 & 8.84\\
6 & 6.08\\
4 & 5.39\\
12 & 8.15\\
7 & 6.42\\
5 & 5.73\\
\bottomrule
\end{tabular}\hfill
\begin{tabular}{rr}
\multicolumn{2}{l}{data\_4}\\
\toprule
x & y\\
\midrule
8 & 6.58\\
8 & 5.76\\
8 & 7.71\\
8 & 8.84\\
8 & 8.47\\
8 & 7.04\\
8 & 5.25\\
19 & 12.50\\
8 & 5.56\\
8 & 7.91\\
8 & 6.89\\
\bottomrule
\end{tabular}



# data

それぞれどういうデータなんだろう

- 表だけ見てもよくわからない

# data

それぞれどういうデータなんだろう

- xの平均
- xの分散
- yの平均
- yの分散
- xとyの相関係数
- 回帰直線の式

# summarise

```{r}
sokan_keisu <- list(df_1, df_2, df_3, df_4) %>% map_dbl(~cor(.x$x, .x$y))
list(df_1, df_2, df_3, df_4) %>% map_dfr(~summarise(.x, across(.cols=everything(), .fns=list(mean=mean, var=var), .names="{col}_{fn}"))) %>% 
  add_column(data=c("data_1", "data_2", "data_3", "data_4"), .before ="x_mean") %>% 
  add_column(cor = sokan_keisu, .after= "y_var") %>% 
  knitr::kable()
```

>- 要約統計量はよく似ている

# visualization

```{r fig.cap="`ggplot2` によるグラフ", echo=FALSE}
anscombe %>% 
  pivot_longer(cols = everything(),
               names_to = c(".value", "set"),
               names_pattern = "(x|y)(\\d)") %>% 
  mutate(set=str_c("data_", set)) %>% 
  ggplot(aes(x, y)) +
  geom_point() +
  facet_wrap(~set)
```



# visualization

```{r fig.cap="`ggplot2` によるグラフ", echo=FALSE}
anscombe %>% 
  pivot_longer(cols = everything(),
               names_to = c(".value", "set"),
               names_pattern = "(x|y)(\\d)") %>% 
  mutate(set=str_c("data_", set)) %>% 
  ggplot(aes(x, y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~set)
```


# lesson

表でみてもよくわからない

平均値とか分散を計算してもよくわからない

図示がだいじ




# research

\LARGE

- \textbullet ある町
- \textbullet\hspace{1pt} 4つの学校
- \textbullet 同じテスト


# research

\LARGE


- 生徒対象に調査
- 　\textbullet テスト勉強にあてた時間
- 　\textbullet 実際の点数

# hypothethis

\LARGE

勉強時間と点数に関係があるでしょうか

あるとしたらどういう関係が\ldots

# plot

```{r echo=FALSE}
df_base <- iris %>%
  janitor::clean_names() %>%
  as_tibble() %>% 
  filter(species == "versicolor") %>% 
  rename("time" = sepal_length,
         "point" = petal_length,
         "school" = species)
```

```{r echo=FALSE}
df_A <- df_base %>% 
  mutate(time = time - 3, point = point + .75, school = "school_A")

df_B <- df_A %>% 
  mutate(time = time -.45, point = point + .75, school = "school_B")

df_C <- df_B %>% 
  mutate(time = time -.45, point = point + .75, school = "school_C")

df_D <- df_C %>% 
  mutate(time = time -.45, point = point + .75, school = "school_D")
```


```{r echo=FALSE}
p <- df_A %>% bind_rows(df_B, df_C, df_D) %>% 
  ggplot(aes(time, point * 10)) + geom_point(alpha = .8) +
  labs(x = "time spent studying", 
       y = "score"
       )
p
```



```{r echo=FALSE}
q <- df_A %>% bind_rows(df_B, df_C, df_D) %>% 
  ggplot(aes(time, point * 10)) + geom_jitter(width = .8, height = .8, alpha = .8) +
  labs(x = "time spent studying", 
       y = "score")
       
q
```


# plot



```{r echo=FALSE, message=FALSE}
q +
  geom_smooth(method = "lm", se = FALSE)
```


# plot

\IfFileExists{./letsnotsee.jpg}{\includegraphics[width=\textwidth]{./letsnotsee.jpg}}{\relax}


# plot



```{r echo=FALSE}
q +
  aes(colour = school) +
  scale_color_viridis_d(name = "")
```




# plot



```{r echo=FALSE}
q +
  aes(colour = school) +
  scale_color_viridis_d(guide = "none") +
  facet_wrap(~school)
```

# plot



```{r echo=FALSE}
q +
  aes(colour = school) +
  scale_color_viridis_d(guide = "none") +
  facet_wrap(~school) +
  geom_smooth(method = "lm", se = FALSE)
```

# plot



```{r echo=FALSE}
q +
  aes(colour = school) +
  scale_color_viridis_d(name = "") +
  geom_smooth(method = "lm", se = FALSE)
```

# plot



```{r echo=FALSE, message=FALSE}
q +
  geom_smooth(method = "lm", se = FALSE)
```



# Simpson's Paradox
\large

-  \textbullet{}\hspace{2pt}全体で見ると「勉強すればするほど成績がさがる」
-  \textbullet{}\hspace{2pt}学校ごとに見ると「勉強すればするほど成績があがる」

\normalsize


# quiz:How many millions are in a trillion?


```{r echo=FALSE}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width,
                      colour = Species)) +
  geom_point()
spc <- c("ヒオウギアヤメ", "イリス・ウェルシコロル",
         "イリス・ヴァージニカ")
p + labs(x = "がくの長さ(cm)", y = "萼の幅(cm)") +
         scale_colour_discrete(name = "種", labels = spc)
```




# quiz:How many millions are in a trillion?


\LARGE

1兆は100万の何倍でしょう？


\begin{tabular}{rr}\toprule
回答&率\\\midrule
千倍&18\%\\
万倍&12\%\\
10万倍&21\%\\
100万倍&21\%\\
1000万倍&17\%\\
わからない&12\%\\\bottomrule
\end{tabular}





# pie chart

```{r include=FALSE}
choices = c("10^3 倍","10^4 倍","10^5 倍","10^6 倍","10^7 倍","わからない") %>% fct_inorder() 
value <-  c(18,12,21,21,17,12)
df_pie <- tibble(choices, value)
```


```{r out.height="8cm"}
df_pie %>% 
　　mutate(position = 100 - (cumsum(value) - value/2)) %>% 
  ggplot(aes(x = "", y = value, fill = choices)) + 
　　geom_bar(stat = "identity", position = "stack") +
　　coord_polar(theta = "y", direction = -1) + 
　　geom_text(aes(y = position, label = choices), size = 6, colour = "white")  +
    scale_fill_viridis_d(guide = "none") +
    theme_void()
```


# substitute for a pie chart


```{r echo=FALSE}
df_pie %>% 
  ggplot() + geom_col(aes(choices, value, fill = choices))  +
  scale_fill_viridis_d(guide = "none") +
  labs(x = "", y = "比率(%)")
```


# substitute for a pie chart


```{r}
df_pie %>% 
  ggplot() + geom_col(aes(fct_rev(choices), value, fill = choices)) +
  scale_fill_viridis_d(guide = "none") +
  coord_flip() +
  labs(x = "", y = "比率(%)")
```


# substitute for a pie chart


```{r}
df_pie %>% 
  ggplot() + geom_point(aes(fct_rev(choices), value, colour = choices),
                        colour = "orange",
                        size = 10) +
  geom_segment(aes(x = choices, xend = choices,
                   y = 0, yend = value),
               colour = "grey",
               size = 1) +
  coord_flip() +
  labs(x = "", y = "比率(%)")
```



# quiz:How many millions are in a trillion?


```{r echo=FALSE}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width,
                      colour = Species)) +
  geom_point()
spc <- c("ヒオウギアヤメ", "イリス・ウェルシコロル",
         "イリス・ヴァージニカ")
p +  scale_colour_discrete(name = "種", labels = iris$Species) +
     labs(x = "ほげほげ(cm)", y = "およよ(cm)") 
```


# Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```


# 文書・資料

To do our work, we all have to read a mass of papers. Nearly all of them are far too long. This wastes time, while energy has to be spent in looking for the essential points.

I ask my colleagues and their staffs to see to it that their Reports are shorter.


# 文書・資料


- \textbullet{}\hspace{2pt} The aim should be Reports which set out the main points in a series of short, crisp paragraphs.
-  \textbullet{}\hspace{2pt} If a Report relies on detailed analysis of some complicated factors, or on statistics, these should be set out in an Appendix.
-  \textbullet{}\hspace{2pt} Often the occasion is best met by submitting not a full-dress Report, but an Aide-memoire consisting of headings only, which can be expanded orally if needed. 
- \textbullet\hspace{2pt}Let us have an end of such phrases as these: 
 “It is also of importance to bear in mind the following considerations...”, or  “Consideration should be given to the possibility of carrying into effect...”
Most of these woolly phrases are mere padding, which can be left out altogether, or replaced by a single word. Let us not shrink from using the short expressive phrase, even if it is conversational.

# 文書・資料


Reports drawn up on the lines I propose may at first seem rough as compared with the flat surface of officialese jargon. But the saving in time will be great, while the discipline of setting out the real points concisely will prove an aid to clearer thinking.





# tohoho


```{r include=FALSE}
choices = c("ほぼ毎日","週2、3回","週1回","2週間に1回","月1回","学校による", "有事のみ", "未実施", "把握せず") %>% fct_inorder() 
value <-  c(11.3,5.7,17,7.5,13.2,3.8,3.8,34,1.9)
df_pie <- tibble(choices, value)
```


```{r}
school <- c("school_A", "school_B", "school_C")
とても不満 <- c(5, 18, 20)
不満 <- c(20, 18, 25)
ふつう <- c(13, 15, 30)
満足 <- c(43, 30, 15)
とても満足 <- c(17, 20, 10)

df_likert <- tibble(school, とても不満, 不満, ふつう, 満足, とても満足)
```

```{r}
df_likert %>%
  pivot_longer(-school) %>%
  mutate(name = fct_inorder(name)) %>% 
  mutate(name = fct_rev(name)) %>% 
  ggplot(data = .)+
  geom_bar(aes(x = fct_rev(school), y=value, fill=name), position="stack", stat="identity")+
  coord_flip() + 
  ggtitle("授業の満足度")+
  ylab("Percentage")+
  xlab("School")+
  scale_fill_brewer(palette="PRGn", name = "" ) +
  theme(legend.position="bottom") +
  guides(fill = guide_legend(reverse = TRUE))
```


# tohoho
 



```{r}
df_likert %>% HH::likert(school~., ., as.percent = TRUE,
           main="授業の満足度",
           xlab="percentage", ylab="School") 
```


